---
title: "Lab 07 Student Problems"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## Student Problem 01

(1) Load the `gad7_data.csv` dataset, which shows repeated observations of the GAD-7 scale in 50 patients, which measures generalized anxiety symptoms.

```{r}

gad_7_data <- read_csv("data/gad7_data.csv")

```

(2) Pivot this data longer so that you can create a spaghetti plot of GAD-7 scores over four weeks. Try to make your spaghetti plot as clear as possible.

```{r}

gad_longer <- gad_7_data %>%
  pivot_longer(cols = starts_with("week"),
               names_to = "week",
               values_to = "gad_score")

ggplot(data = gad_longer) +
  geom_line(aes(x = week, y = gad_score, group = id), alpha = 0.4) +
  theme_bw()

# You can manually fix the week variable names, but this wouldn't scale
# well if you had a lot of weeks
gad_longer_f <- gad_longer %>%
  mutate(week = case_when(week == "week1_gad" ~ 1,
                          week == "week2_gad" ~ 2,
                          week == "week3_gad" ~ 3,
                          week == "week4_gad" ~ 4))

ggplot(data = gad_longer_f) +
  geom_line(aes(x = week, y = gad_score, group = id), alpha = 0.4) +
  theme_bw()

# Within pivot use the argument names_transform to change names
gad_longer_nt <- gad_7_data %>%
  pivot_longer(cols = starts_with("week"),
               names_to = "week",
               values_to = "gad_score",
               names_transform = list("week" = function(x) str_remove(x, "week")))

# Write a quick cleaning function that alters the names and turns them into 
# numerics
cleaning_function <- function(s){
  step1 <- str_remove(s, "week") 
  step2 <- str_remove(step1, "_gad")
  step3 <- as.numeric(step2)
  return(step3)
}

# Call your cleaning function inside of the names_transform argument
gad_longer_nt <- gad_7_data %>%
  pivot_longer(cols = starts_with("week"),
               names_to = "week",
               values_to = "gad_score",
               names_transform = list("week" = function(x) cleaning_function(x)))

# Able to write your function directly into names_transform argument:
gad_longer_nt_alt <- gad_7_data %>%
  pivot_longer(cols = starts_with("week"),
               names_to = "week",
               values_to = "gad_score",
               names_transform = list("week" = function(x){
                  step1 <- str_remove(x, "week") 
                  step2 <- str_remove(step1, "_gad")
                  step3 <- as.numeric(step2)
                  return(step3)
                }))

```

(3) Create an indicator variable to identify individuals whose GAD score increases from Week 3 to Week 4. As a first step, pivot_wider() from the long dataset to reproduce the original dataset.

```{r}

# Pivoting wide makes it much easier to do this:
gad_wider <- gad_longer_nt_alt %>%
  pivot_wider(names_from = week,
              values_from = gad_score,
              names_prefix = "gad_week_")

# Just compare week 4 and week 3 variables -- easy!
gad_mutated <- gad_wider %>%
  mutate(late_increasers = if_else(gad_week_4 > gad_week_3, 1, 0))

```

(4) Now create an indicator variable to identify individuals whose GAD score decreases every week throughout the study.

```{r}

gad_mutated_more <- gad_mutated %>%
  mutate(decreasers = if_else(gad_week_4 < gad_week_3 &
           gad_week_3 < gad_week_2 &
           gad_week_2 < gad_week_1, 1, 0))


```

(5) Create a variable for those who don't fall into the two above categories -- call this variable "others".

```{r}

gad_mutated_final <- gad_mutated_more %>%
  mutate(others = if_else(late_increasers == 0 & decreasers == 0, 1, 0))

```

(6) Create a faceted spaghetti plot by these three new categories that you just created.

```{r}

# Strategy 1: pivot longer twice

gad_mutated_longer1 <- gad_mutated_final %>%
  pivot_longer(cols = starts_with("gad_week"),
               names_to = "week",
               values_to = "gad_score",
               names_transform = list("week" = function(x){
                 s1 <- str_remove(x, "gad_week_")
                 out <- as.numeric(s1)
                 return(out)
               }))

gad_mutated_longer2 <- gad_mutated_longer1 %>%
  pivot_longer(cols = late_increasers:others,
               names_to = "class",
               values_to = "presence") %>%
  filter(presence == 1)

# Strategy 2 pivot longer and then create a new variable:

gad_mutated_longer1 <- gad_mutated_final %>%
  pivot_longer(cols = starts_with("gad_week"),
               names_to = "week",
               values_to = "gad_score",
               names_transform = list("week" = function(x){
                 s1 <- str_remove(x, "gad_week_")
                 out <- as.numeric(s1)
                 return(out)
               }))

gad_mutated_longer_alt <- gad_mutated_longer1 %>%
  mutate(class = case_when(late_increasers == 1 ~ "Late Increaser",
                           decreasers == 1 ~ "Decreaser",
                           others == 1 ~ "Other"))

ggplot(data = gad_mutated_longer_alt) +
  geom_line(aes(x = week, y = gad_score, group = id), alpha = 0.4) +
  theme_bw() +
  facet_wrap(~class)



```

## Student Problem 02

(1) Load the latest NYC COVID-19 data from https://raw.githubusercontent.com/nychealth/coronavirus-data/master/latest/now-cases-by-day.csv and clean the dates using lubridate::mdy

```{r}

covid_data <- read_csv("https://raw.githubusercontent.com/nychealth/coronavirus-data/master/latest/now-cases-by-day.csv")

clean_covid <- covid_data %>%
  mutate(clean_date = lubridate::mdy(date_of_interest)) # this is a way to call a function from a package you have installed without loading the whole library

```

(2) Plot the total case count (`CASE_COUNT`) over time.

```{r}

ggplot(data = clean_covid) +
  geom_line(aes(x = clean_date, y = CASE_COUNT)) +
  theme_bw()


```

(3) Use `pivot_longer()` to create a dataset that will allow you to map different borough case counts to the color aesthetic `BX_CASE_COUNT`, `BK_CASE_COUNT`, `MN_CASE_COUNT`, `SI_CASE_COUNT`, `QN_CASE_COUNT`.

```{r}

# Pivot these multiple columns into a variable that represents borough and a variable
# that represents cases
borough_long <- clean_covid %>%
  pivot_longer(cols = c("BX_CASE_COUNT",
                        "BK_CASE_COUNT",
                        "MN_CASE_COUNT",
                        "QN_CASE_COUNT",
                        "SI_CASE_COUNT"),
               names_to = "borough",
               values_to = "cases")

# Now it is very easy to create graph with different boroughs mapped onto
# the color aesthetic
ggplot(data = borough_long) +
  geom_line(aes(x = clean_date, y = cases, color = borough)) +
  theme_minimal()

# geom_ribbon creates a better visualization of these multiple lines
ggplot(data = borough_long) +
  geom_ribbon(aes(x = clean_date, ymin = 0, ymax = cases, fill = borough), alpha = 0.4) +
  theme_minimal()

```

(4) Create a similar plot of COVID-19 probable case counts and then place these two plots in a vertical column using the **patchwork** package.

```{r}


# Reproducing the borough case count long dataset from above
borough_long_case <- clean_covid %>%
  pivot_longer(cols = c("BX_CASE_COUNT",
                        "BK_CASE_COUNT",
                        "MN_CASE_COUNT",
                        "QN_CASE_COUNT",
                        "SI_CASE_COUNT"),
               names_to = "borough",
               values_to = "cases")

# Creating the plot from above, but saving it as an object in our environment so
# we can manipulate it
case_plot <- ggplot(data = borough_long_case) +
  geom_line(aes(x = clean_date, y = cases, color = borough)) +
  theme_minimal()

# Now go through the same process for the probable case counts
borough_long_prob <- clean_covid %>%
  pivot_longer(cols = c("BX_PROBABLE_CASE_COUNT",
                        "BK_PROBABLE_CASE_COUNT",
                        "MN_PROBABLE_CASE_COUNT",
                        "QN_PROBABLE_CASE_COUNT",
                        "SI_PROBABLE_CASE_COUNT"),
               names_to = "borough",
               values_to = "prob_cases")

prob_plot <- ggplot(data = borough_long_prob) +
  geom_line(aes(x = clean_date, y = prob_cases, color = borough)) +
  theme_minimal()

# Using the patchwork package you can combine plots that are saved in your environment
library(patchwork)
# Install if necessary:
#install.packages("patchwork") 

# Plots side by side using | operator
case_plot | prob_plot

# Plots on top vertically using / operator
case_plot / prob_plot

# You can combine in multiple ways:
(case_plot / prob_plot) | prob_plot


```

Check out more details in the patchwork documentation:

(https://patchwork.data-imaginist.com/articles/patchwork.html)


(5) Create a faceted plot (case/probable case) instead of the two separate plots in (4).

```{r}

# Select down to make these manageable, recode the boroughs so they are the same for joining
borough_cases <- borough_long %>%
  select(clean_date, borough, cases) %>%
  mutate(borough = str_remove(borough, "_CASE_COUNT"))

borough_probable <- borough_long_prob %>%
  select(clean_date, borough, prob_cases) %>%
  mutate(borough = str_remove(borough, "_PROBABLE_CASE_COUNT"))

# Join together

borough_all <- borough_cases %>%
  full_join(borough_probable, by = c("clean_date", "borough"))

all_long <- borough_all %>%
  pivot_longer(cols = cases:prob_cases,
               names_to = "case_type",
               values_to = "count")

# Now we have a dataset where you can facet by type of plot:
ggplot(data = all_long) +
  geom_line(aes(x = clean_date, y = count, color = borough)) +
  geom_ribbon(aes(x = clean_date, ymin = 0, ymax = count, fill = borough), alpha = 0.4) +
  facet_wrap(~case_type, ncol = 1) +
  theme_bw()



```
